The Soulgem Paper

### 2011-09-22 soulgem.tex
added some packages
A few more paragraphs in the Findings section

@@

@@

In summary, as proposed by our theoretical framework, \textbf{we found a considerable inconsistency in role definitions both within and between organizations}. Roles (sets of expectations) are socially constructed: a developer here is not the same as a developer there, since what we expect of developers here and there varies. Part of the social construction occurs at the societal level, so that we generally expect testers to test, developers to develop, and managers to manage. In one case, such as that of Project Managers at Gallium, this broader influence is strong enough to shape most aspects of the role. In most roles in the software organizations we studied, however, this is not the case, and most of the relevant details of what it actually means to test, develop, or manage is defined by the group.

% Common roles? common expectations?

>>>>

\usepackage{comment}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{array}
\usepackage{booktabs}
\usepackage{capt-of}
\usepackage{colortbl}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfig}
%\DeclareCaptionType{copyrightbox}
\usepackage{pifont}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{url}
\usepackage{boxedminipage}
\usepackage{xspace}
\usepackage{sepnum}
\usepackage{cite}
\usepackage{fancyhdr}

@@

\input{expectationPeopleTable}

@@

In summary, as proposed by our theoretical framework, \textbf{we found inconsistent role definitions both within and between organizations}. Roles (sets of expectations) are socially constructed: a developer here is not the same as a developer there, since what we expect of developers here and there varies. Part of the social construction occurs at the societal level, so that we generally expect testers to test, developers to develop, and managers to manage. In one case, such as that of Project Managers at Gallium, this broader influence is strong enough to shape most aspects of the role. In most roles in the software organizations we studied, however, this is not the case, and most of the relevant details of what it actually means to test, develop, or manage is defined by the group.

\subsection{Role Commonalities and Expectations (RQ2)}

As expected, we could not provide a simple classification of roles. However, we believe that, should we continue studying more organizations, some patterns in role definitions would begin to emerge.

We can see the beginnings of such patterns in Table X, which shows the expectations placed on the people we interviewed (but not on all the members of the organizations we studied). The columns in the table are arranged by positions, so that people with roughly similar positions\footnote{An exact matching of people to positions was impractical.} in the two organizations are grouped together. The rows in the table represent common or significant expectations. The reader should note that this list of expectations is necessarily incomplete for the general case and somewhat subjective, as expectations in our interviews and observations were not always clearly articulated and our participants did not always agree on their expectation assignments.

In the case of experienced developers, for instance, we found a shifting emphasis towards providing guidance to newcomers, helping team members understand the rationale behind design decisions in legacy code, experimenting with novel approaches and techniques, having a greater say in architectural decisions, and providing the organization with technical authority in its meetings with current or potential customers. In other words, they transition towards expectations that are centred on their experience and expertise.

% TODO: Revisit this. Do we have something to say about other role flavours here?

\subsection{Expectation Clustering (RQ3)}

We proposed that expectations are clustered in roles based on a combination of institutional, historical, structural, and personal factors. Our empirical evidence provided support for this proposition, as detailed below.

\emph{Institutional factors:} The larger culture plays an important part in shaping the expectations of members of software organizations. For example, when Gallium hired one of its Product Managers, he had significant domain expertise, but no software development experience. He initially struggled to satisfy the expectations placed on him. (...)

\emph{Historical factors:}

\emph{Structural factors:}

\emph{Personal factors:}

\subsection{Divergence between Roles and Positions (RQ4)}
